---
title: "Assignment 3"
author: "Xingyu Wang"
date: "2023-10-13"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Exploratory data analysis
```{r}
library(tidyverse)
library(Stat2Data)
data("Hawks")
```

## 1.1 (Q1) 
```{r}
# Creating the HawksTail vector
HawksTail <- Hawks$Tail

# Display the first few elements of the vector
head(HawksTail) 
# Output: [1] 219 221 235 220 157 230

# Compute the sample mean and sample median
sample_mean <- mean(HawksTail, na.rm = TRUE)  # using na.rm = TRUE to handle any NA values
sample_median <- median(HawksTail, na.rm = TRUE) 

# Display the sample mean and sample median
sample_mean
sample_median

```


## 1.2 (Q1)
```{r}
# Using summarise() to compute mean, trimmed mean, and median for Wing and Weight columns
result <- Hawks %>%
  summarise(
    Wing_mean = mean(Wing, na.rm = TRUE),
    Wing_t_mean = mean(Wing, trim = 0.5, na.rm = TRUE),  # trimmed mean with q=0.5
    Wing_med = median(Wing, na.rm = TRUE),
    Weight_mean = mean(Weight, na.rm = TRUE),
    Weight_t_mean = mean(Weight, trim = 0.5, na.rm = TRUE),  # trimmed mean with q=0.5
    Weight_med = median(Weight, na.rm = TRUE)
  )

# Display the result
print(result)

```
## 1.2 (Q2)
```{r}

# Grouping by Species and then using summarise() to compute mean, trimmed mean, and median
grouped_result <- Hawks %>%
  group_by(Species) %>%
  summarise(
    Wing_mean = mean(Wing, na.rm = TRUE),
    Wing_t_mean = mean(Wing, trim = 0.5, na.rm = TRUE),  # trimmed mean with q=0.5
    Wing_med = median(Wing, na.rm = TRUE),
    Weight_mean = mean(Weight, na.rm = TRUE),
    Weight_t_mean = mean(Weight, trim = 0.5, na.rm = TRUE),  # trimmed mean with q=0.5
    Weight_med = median(Weight, na.rm = TRUE)
  )

# Display the grouped result
print(grouped_result)

```
....

## 1.3 (Q1)
<div style="height: 200px;"></div>












```{r}
a <- 2
b <- 3

transformed_mean <- mean(HawksTail * a + b)
calculated_mean <- a * mean(HawksTail) + b

# Compare the two means:
transformed_mean
calculated_mean


```

## 1,3 (Q2)
<div style="height: 200px;"></div>









```{r}
transformed_variance <- var(HawksTail * a + b)
calculated_variance <- a^2 * var(HawksTail)

transformed_sd <- sd(HawksTail * a + b)
calculated_sd <- a * sd(HawksTail)

# Compare the variance and standard deviation:
transformed_variance
calculated_variance

transformed_sd
calculated_sd

```
....

## 1.4
```{r}
hal<-Hawks$Hallux # Extract the vector of hallux lengths
hal<-hal[!is.na(hal)] # Remove any nans
outlier_val<-100
num_outliers<-10
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
mean(hal)

mean(corrupted_hal)

num_outliers_vect <- seq(0,1000)
means_vect <- c()
for(num_outliers in num_outliers_vect){
corrupted_hal <- c(hal,rep(outlier_val,times=num_outliers))
means_vect <- c(means_vect, mean(corrupted_hal))
}

```

## 1.4 (Q1)
```{r}
num_outliers_vect <- seq(0,1000)
medians_vect <- c()

for(num_outliers in num_outliers_vect){
  corrupted_hal <- c(hal, rep(outlier_val, times=num_outliers))
  medians_vect <- c(medians_vect, median(corrupted_hal))
}

```
## 1.4 (Q2)
```{r}
num_outliers_vect <- seq(0,1000)
t_means_vect <- c()

for(num_outliers in num_outliers_vect){
  corrupted_hal <- c(hal, rep(outlier_val, times=num_outliers))
  t_means_vect <- c(t_means_vect, mean(corrupted_hal, trim = 0.1))
}

```
## 1.4 (Q3)
```{r}
# Combining into a dataframe
df_means_medians <- data.frame(
  num_outliers=num_outliers_vect, 
  mean=means_vect,
  t_mean=t_means_vect, 
  median=medians_vect
)

# Reshape data for plotting
df_long <- df_means_medians %>% 
  pivot_longer(
    cols = -num_outliers, 
    names_to = "statistic", 
    values_to = "value"
  )

# Plotting
ggplot(df_long, aes(x=num_outliers, y=value, color=statistic)) + 
  geom_line() + 
  labs(title="Effect of Outliers on Location Estimators", 
       x="# of Outliers", 
       y="Value of Location Estimator") +
  theme_minimal()

```

## 1.5(Q1)
```{r}
# Plotting the data
ggplot(Hawks, aes(x=Species, y=Weight)) + 
  geom_boxplot(outlier.color = "black", outlier.shape = 16) +
  labs(title="Distribution of Hawk Weights by Species", 
       x="Species", 
       y="Weight (g)") +
  theme_minimal()

```
## 1.5(Q2)
```{r}
# Grouping by species and computing quantiles
quantiles_df <- Hawks %>%
  group_by(Species) %>%
  summarise(
    quantile025 = quantile(Weight, 0.25, na.rm = TRUE),
    quantile050 = quantile(Weight, 0.50, na.rm = TRUE),
    quantile075 = quantile(Weight, 0.75, na.rm = TRUE)
  )

print(quantiles_df)

```
quantile025 corresponds to the lower hinge (the bottom of the box) of the boxplot.
quantile050 corresponds to the median (the line inside the box) of the boxplot.
quantile075 corresponds to the upper hinge (the top of the box) of the boxplot.
## 1.5(Q3)
```{r}
num_outliers <- function(sample_vector) {
  # Removing NA values
  sample_vector <- sample_vector[!is.na(sample_vector)]
  
  # Calculating the quantiles
  q25 <- quantile(sample_vector, 0.25)
  q75 <- quantile(sample_vector, 0.75)
  
  # Calculating the Interquartile Range (IQR)
  IQR <- q75 - q25
  
  # Finding outliers based on the provided conditions
  lower_bound <- q25 - 1.5 * IQR
  upper_bound <- q75 + 1.5 * IQR
  
  outliers <- sample_vector[sample_vector < lower_bound | sample_vector > upper_bound]
  
  return(length(outliers))
}

# Testing the function
num_outliers( c(0, 40, 60, 185))

```
## 1.5(Q4)
```{r}
outliers_by_species <- Hawks %>%
  group_by(Species) %>%
  summarise(
    num_of_outliers = num_outliers(Weight)
  )

print(outliers_by_species)

```
## 1.6(Q1)
```{r}
# Compute covariance
cov_weight_wing <- cov(Hawks$Weight, Hawks$Wing, use = "complete.obs")

# Compute correlation
cor_weight_wing <- cor(Hawks$Weight, Hawks$Wing, use = "complete.obs")

cat("Covariance between Weight and Wing:", cov_weight_wing, "\n")
cat("Correlation between Weight and Wing:", cor_weight_wing, "\n")

```
## 1.6(Q2)
```{r}
# Assuming you have a dataframe Hawks with columns Weight and Wing

a <- 2.4
b <- 7.1
c <- -1
d <- 3

# Creating the transformed variables
Hawks$Weight_transformed <- a * Hawks$Weight + b
Hawks$Wing_transformed <- c * Hawks$Wing + d

# Computing covariance and correlation for transformed variables
cov_transformed <- cov(Hawks$Weight_transformed, Hawks$Wing_transformed, use = "complete.obs")
cor_transformed <- cor(Hawks$Weight_transformed, Hawks$Wing_transformed, use = "complete.obs")

# Displaying the results
cat("Covariance of transformed variables:", cov_transformed, "\n")
cat("Correlation of transformed variables:", cor_transformed, "\n")

```
## 2.1(Q1)
1. Random Experiment: A random experiment is an experiment or a process for which the outcome cannot be predicted with certainty.
2. Sample Space:
The sample space, often denoted as S or Ω, refers to the set of all possible outcomes of a random experiment. It encompasses every conceivable result for the given experiment
3. Event: An event is any subset of the sample space. It represents a specific set of outcomes of a random experiment that we might be interested in. 

## 2.1 (Q2)
<div style="height: 200px;"></div>











## 2.2 (Q1)
<div style="height: 200px;"></div>

## 2.2 (Q2)
<div style="height: 200px;"></div>


## 2.2 (Q2)
<div style="height: 200px;"></div>

## 2.2 (Q3)
<div style="height: 200px;"></div>

## 2.2 (Q4)
<div style="height: 200px;"></div>

## 2.2 (Q5)
<div style="height: 200px;"></div>


## 3 (Q1)
```{r}
# Load the required libraries
library(ggplot2)

# Create the density plot
ggplot(data = Hawks, aes(x = Tail, fill = Species)) + 
  geom_density(alpha = 0.5) + 
  labs(title = "Density plot of Hawk Tail Lengths by Species",
       x = "Tail (mm)",
       y = "Density") +
  theme_minimal()


```
## 3 (Q2)
```{r}
# Load the required libraries
library(ggplot2)

# Create the violin plot
ggplot(data = Hawks, aes(x = Species, y = Tail, fill = Species)) + 
  geom_violin() + 
  labs(title = "Violin plot of Hawk Tail Lengths",
       x = "Species",
       y = "Tail (mm)") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1")

```
## 3 (Q3)
```{r}
# Define the custom shapes and colors for species
hawks_shapes <- c(CH = 16, RT = 17, SS = 15)  # 16: circle, 17: triangle, 15: square
hawks_colors <- c(CH = "red", RT = "green", SS = "blue")

# Plot
ggplot(Hawks, aes(x = Tail, y = Weight, shape = Species, color = Species)) +
  geom_point(aes(shape = Species, color = Species), size = 3) +
  scale_shape_manual(values = hawks_shapes) +
  scale_color_manual(values = hawks_colors) +
  labs(title = "Tail vs. Weight for different Hawks species",
       x = "Tail (mm)",
       y = "Weight",
       shape = "Species",
       color = "Species")


```
## 3 (Q4)
```{r}

# Define the custom shapes and colors for species
hawks_shapes <- c(CH = 16, RT = 17, SS = 15)  # 16: circle, 17: triangle, 15: square
hawks_colors <- c(CH = "red", RT = "green", SS = "blue")

# Plot
ggplot(Hawks, aes(x = Tail, y = Weight, shape = Species, color = Species)) +
  geom_point(aes(shape = Species, color = Species), size = 3) +
  geom_smooth(method = "lm", se = TRUE, aes(color = Species), show.legend = FALSE) + 
  scale_shape_manual(values = hawks_shapes) +
  scale_color_manual(values = hawks_colors) +
  facet_wrap(~ Species, scales = "free") +
  labs(title = "Tail vs. Weight for different Hawks species",
       x = "Tail (mm)",
       y = "Weight",
       shape = "Species",
       color = "Species")

```
.....
## 3 (Q5)
```{r}

# Find the heaviest hawk
heaviest_hawk <- Hawks %>%
  filter(Weight == max(Weight, na.rm = TRUE)) %>%
  select(Tail, Weight, Species) %>%
  top_n(1)

tail_heaviest <- heaviest_hawk$Tail
weight_heaviest <- heaviest_hawk$Weight

# Define the custom shapes and colors for species
hawks_shapes <- c(CH = 16, RT = 17, SS = 15)  # 16: circle, 17: triangle, 15: square
hawks_colors <- c(CH = "red", RT = "green", SS = "blue")

# Plot
ggplot(Hawks, aes(x = Tail, y = Weight, shape = Species, color = Species)) +
  geom_point(aes(shape = Species, color = Species), size = 3) +
  scale_shape_manual(values = hawks_shapes) +
  scale_color_manual(values = hawks_colors) +
  labs(title = "Tail vs. Weight for different Hawks species",
       x = "Tail (mm)",
       y = "Weight",
       shape = "Species",
       color = "Species") +
  geom_segment(aes(x = tail_heaviest, xend = tail_heaviest + 10, 
                   y = weight_heaviest, yend = weight_heaviest + 100), 
               arrow = arrow(type = "closed", length = unit(0.2, "inches")), color = "black") +
  annotate("text", x = tail_heaviest + 25, y = weight_heaviest + 150, 
           label = "heaviest hawk", color = "black")


```


